Report

### 2022-07-29 manuscript.Rmd
Start implementation of conditions into discussion

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators. 
Each of these approaches requires making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

@@

This tutorial paper introduces the first implementation of BES in user-friendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

The main algorithm of interest was the PBF.
As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was vote counting: counting the number of significant effects.
Although this method is in use for aggregating conceptual replications,
it should not be considered good practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [(Scheerens, 2015):].
Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extention, $H_i$).
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
The third algorithm was individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.
<!-- Rebecca: -->
<!-- zou je dit vooraf moeten zeggen? -->
<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->

@@

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of reliability on the diference between algorithms (see Table \@ref(tab:tabeffect)).

@@

Within the scope of this simulation, PBF had relatively better inferential properties than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
<!-- Conventional threshold for type 1 error: 5% -->
<!-- Conventional threshold for power: 80% -->
<!-- J. Cohen, Statistical Power Analysis for the -->
<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
<!--  NJ, 1988). This is the source of the system of power -->
<!--  analysis described here; the power values and sam -->
<!--  ple sizes of the illustrations derive from this book's -->
<!--  tables. -->
<!--  2. J -->

An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
Overall, PBF had the best performance, with power of 76% and type I error of 24%.
Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations should be acknowledged in applied research.

The overall performance of PBF increased most with increasing sample size.
With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
With increasing reliability, increasing specificity was traded off for decreasing sensitivity.

@@

<!-- Hier ook dan artikel van BEM noemen, maar oook zeggen dat het daar om null hypotheses gaat. -->

>>>>

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators. 
Each of these approaches require making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

@@

This tutorial paper introduces the first implementation of BES in useruse-friendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

@@

The main algorithm of interest was the PBF.
As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was vote counting: counting the number of significant effects.
Although this method is in use for aggregating conceptual replications,
it should not be considered good practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [(Scheerens, 2015):].
Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
The third algorithm was individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.
<!-- Rebecca: -->
<!-- zou je dit vooraf moeten zeggen? -->
<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->

@@

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of reliability on the difference between algorithms (see Table \@ref(tab:tabeffect)).

@@

This simulation study validated the product Bayes factor as an alternative method to synthesize evidence across studies.
The PBF was benchmarked against other common methods for evidence synthesis.
This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.

Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.

An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
Overall, PBF had the best performance, with a power of 76% and type I error of 24%.
Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations <!-- higher probability for type 1 errors --> should be acknowledged in applied research.

<!-- Conventional threshold for type 1 error: 5% -->
<!-- Conventional threshold for power: 80% -->
<!-- J. Cohen, Statistical Power Analysis for the -->
<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
<!--  NJ, 1988). This is the source of the system of power -->
<!--  analysis described here; the power values and sam -->
<!--  ple sizes of the illustrations derive from this book's -->
<!--  tables. -->
<!--  2. J -->

<!-- Hier ook dan artikel van BEM noemen, maar ook zeggen dat het daar om null hypotheses gaat. -->

The overall performance of PBF increased most with increasing sample size.
PBF's sensitivity was .57 when a sample size of 20 per sample was simulated, while for all other algorithms this was below .11.
Even at a sample size of 500, the sensitivity of the other algorithms only rose slightly above .60, 
while for PBF this was .90.
This provides evidence that state-of-the-art evidence synthesis methods have little power to detect a true effect.
PBF on the other hand presents itself as a small sample size solution for evidence synthesis methods to detect true effects.
This is reflected in the accuracy of the algorithms at small sample sizes.
The accuracy at a sample size of 20, was around .50 for RMA, IPD and VC.
This can be attributed to their low sensitivity and their specificity that is at ceiling level. 
By consistently not finding an effect, the algorithms are correct 50\% of the time.
At all sample sizes, the PBF was the most accurate algorithm,
although it should be noted that even at a sample size of 500, PBFs specificity was .12 percentage points too low for conventional acceptable type I error rates.

With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!-- I would argue that the decrease is negligable, 0.767 at k =2, 0.763 at k = 10 -->
With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
Again, the specificity remained at ceiling level for the other algorithms.
The sensitivity for RMA and IPD increased, but stayed below .51.
In a similar way, this supports the use of the PBF as a small sample size solution.
In this case when there are little available studies to synthesize evidence over.
Even with as much as ten available studies, PBF was the only algorithm that has a conventionally acceptable power.

With increasing reliability, increasing specificity was traded off for decreasing sensitivity.

There was no marginal condition in which the PBF had a conventional acceptable type I error rate.
Its mean was .76 which is almost .20 percentage points below .95.
However, the other algorithms had no marginal condition where an acceptable power was achieved.
The PBF can be used as a small sample size solution,
although its increased type I error rate must be taken into account.
It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.

@@
